---
title: "Using Panel Maps to Transform Data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using Panel Maps to Transform Data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(conformr)
library(dplyr)
```

# Transformation Steps

Assuming you have a valid panel map, how can you use this to transform a single data series?

```{r basic-setup}
## TODO: write panel_map with four cases

## TODO: write toy data-set

```

First up, you should check your panel map covers all codes present in your dataset.

```{r coverage-check}
## TODO: code demo for coverage-check

```

<!-- CH: should the coverage step generate placeholder instruction for missing categories?-->

```r
## TODO: implement check_coverage() with these tests:
dropA_pm <- good_pm[good_pm$std_A != "x1111",] ## remove code instruction
## check_cd_coverage ----
test_that("incomplete `code_dict` raises check_cd_coverage() warning", {
  expect_warning(check_cd_coverage(data = data_in, code_dict = dropA_dict, code_in = "std_A"))
})
```


Next, consider what you want to do with missing values in your data series. For now, we will convert them all to zeroes.

```{r na-to-zero}
## TODO: convert NA to zero in data
```

Next, join your panel map and data together. Let's call the joined table a **data_map**, because it contains both the data to be converted, and the transformation mapping:

```{r data-map}
## TODO: data-map using dplyr::left_join()
```

```{=html}
<!-- What is best practice? 
maybe : `left_join(x = data_in, y = panel_map)`?
-   ignores any extra code_in in the code map that don't appear in data_in
-   doesn't lose any data if some code_in is missing from code map -- just fills in NA for code_out
-->
```
Finally, transform and aggregate the data by their new classification codes:

```{r aggregate-data-map}
## TODO: code demo of aggregation step
```

## Using {conformr} tools

For convenience, the package provides the wrapper function `use_panel_map()`, which will:

-   join your data and `panel_map` by a shared `code_in` column
-   summarise by the output classification, retaining as many `NA` values as possible. <!-- add link to bottom section? -->
-   retain any grouping variables in the source `data`

but does (not?) check:

-   complete coverage of data by the `panel_map`

```{=html}
<!--
CH: add note on coverage 
// provide check_coverage() as a utility,
// have use_panel_map() warn if any code/value in rows couldn't be matched
-->
```

```{r use_panel_map}
code_in <- sym("std_A")

# get data
data_in <- conformr:::toy_AB$data_in %>%
  select(country, 
         std_A,
         A_100,
         A_prod)

# make panel_map
code_dict <- conformr:::toy_AB$codes_BA
toy_pm <- code_dict %>% 
  make_panel_map_equal(.,  code_in = !!code_in, code_out = std_B, "split_A")

# use panel_map
data_out <-
  use_panel_map(map = toy_pm,
                data = group_by(data_in, country), values_from = A_100,
                from_code = std_A, to_code = std_B,
                weights = split_A, .suffix = "_out")
```
## Multi-Variable Transformations

What if you want to use the same panel map on multiple numeric variables?

For example, you have a table with both nominal and real dollar values, and want to apply the same concordance to both series.

One simple solution that might come to mind is splitting your original table into two tables, one for each variable, and then using the panel map separately on both, before finally recombining into a single table. This solution is functional, though perhaps a little clunky as the number of variables increases.

A more elegant solution would be to pivot your data into long format, apply the transformation, and then (optionally) pivot back to wide format.

```{r, eval=FALSE}
## TODO: add toy example for pivot long, then transform

data %>%
    dplyr::select(code_from, {{values_from}}) %>%
  # pivot_longer to turn values_from into single column
    tidyr::pivot_longer(., {{values_from}}, names_to = "from_name", values_to = "value_in")

```

Again, `use_panel_map()` handles the pivoting for you:
```{r}
use_panel_map(map = toy_pm,
              data = group_by(data_in, country), 
              values_from = c(A_100, A_prod),
              from_code = std_A, to_code = std_B,
              weights = split_A, .suffix = "_out")
```


## Grouped Transformations

<!--# TODO: Example of how to use multi-column key instead of "code_in"; e.g. unique panel map for each country or year? -->

<!--# CH: related to grouped panel_maps? -->

  # http://svmiller.com/blog/2021/02/gank-world-bank-data-with-wdi-in-r/

```{r grouped-transformation}
# write sample data
gdp <- 

# write panel_map
tribble(~year, ~country_in, ~country_out,
        
        )


```

## Multi-step Transformations

Example: ISIC4 -> ISIC3 -> ISIC2
-   be careful with handling of column names..

<!--# TODO: write motivation/example -->

<!--# TODO: include demo code -->
```{r multi-step}
# first conversion step


# second conversion step
```

# Usage Notes

## Coverage Checks

You should always check the coverage of panel map for data you want to transform.

<!--# TODO: insert demo of coverage check fnc -->

## Dealing with Missing Values at Aggregation

Recommend dealing with missingness in source data, BEFORE trying to convert it between classifications.

<!--# TODO: explain the issue, and how conformr handles it -->

-   doesn't matter at the "transformation" stage, but doesn't matter for aggregation stage for code_out

<!-- CH: need to add more detail about what each stage is, and how missingness impacts dataset creation -->

### Hierarchy of Missingness

It is generally not possible to fully retain missing values when transforming data from one classification into another. Consider the following transformation with both a many-to-one and one-to-one mapping:

```{r fruit-NA}
# example panel map
fruit_pm <- tribble(~old, ~new, ~weight,
                    "apples", "fruit", 1,
                    "bananas", "fruit", 1,
                    "oranges", "fruit", 1,
                    "carrot", "veg", 1,)

# example data with NA
fruit_data <- tribble(~product, ~sales,
                      "apples", 3250,
                      "bananas", NA,
                      "oranges", 750,
                      "carrot", NA)

# join panel map and data
data_map <- fruit_pm %>%
  rename(product = old) %>%
  full_join(x = ., 
            y = fruit_data,
            by = "product") %>%
  mutate(sales_new = weight * sales) %>%
  rename(category = new)
```

Let's continue the concordance and try preserving all `NA` values from the source `fruit_data`:

```{r fruit-NA-preserve}
# "preserve" missingness?
data_map %>%
  group_by(category) %>%
  summarise(sales = sum(sales_new, na.rm=FALSE))
```

Notice that we lose the data for `apples` and `oranges` because in R `NA + NA = NA`.

We could instead exclude `NA` values from being summed, which is equivalent to converting all missing values to zeros before aggregating:

```{r fruit-NA-convert}
# throw away NAs na.rm=TRUE
data_map %>%
  group_by(category) %>%
  summarise(sales = sum(sales_new, na.rm=TRUE))
```

This approach is likely fine in cases where `NA` and 0 are interchangeable.

However, there might be cases where we want to be greedy with our `NA` values, only giving them up when necessary to prevent loss of actual data. In this case, `conformr` provides the `sum_NA()` function, which powers the aggregation step inside `use_panel_map()`:

```{r fruit-NA-greedy}
# preserve NA 
data_map %>%
  group_by(category) %>%
  summarise(sales = sum_greedy(sales_new))
```
