---
title: "Convert data between classification standards"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{convert}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval=FALSE
)
```

# Use Cases
* labour/trade classifications
* country-year panels

# Demonstrated Workflows
* creating code-dict from data in target standard -- // no match category?
* verifying valid/complete code-dict / concordance mapping (`code-dict-tools` family)
* multi-step conversion pipeline (e.g. ISIC4-3.1-2)

# Demo dataset
* comtrade (2012-2007, for two countries, import)
```{r data-retrieval}
library(comtradr)

q <- list()

q$Germany <- comtradr::ct_search(reporters = "Australia", 
               partners = c("Germany"), 
               start_date = 2016,
               end_date = 2017,
               type = "goods",
               trade_direction = "imports",
               commod_codes = c("All"))

q$Indonesia <- comtradr::ct_search(reporters = "Australia", 
               partners = c("Indonesia"), 
               start_date = 2016,
               end_date = 2017,
               type = "goods",
               trade_direction = "imports",
               commod_codes = "All")
```

```{r}
monthly <- ct_search(reporters = "Australia", 
               partners = c("Indonesia"), 
               start_date = 2016,
               end_date = 2016,
               freq = "monthly",
               type = "goods",
               trade_direction = "imports",
               commod_codes = "Total")
```


```{r}
library(dplyr)
q$Indonesia %>% glimpse()

q$Indonesia %>%
  filter(aggregate_level == 4, trade_flow == "Import") %>%
  select(classification, period, reporter_iso, partner_iso, commodity_code, trade_value_usd)
```


```{r get-concordance, eval=FALSE}
concordance::hs5_hs4 %>% 
  select(ends_with("_4d")) %>%
  distinct() %>%
  conformr::make_panel_map_equal(., HS5_4d, HS4_4d) %>%
  filter(n_dest != 1)
```


```{r setup}
library(dplyr)
# TODO: highlight code, value 2D transformation.
```

## make code-dict (equal weights)
```
## from codes
countrycode::<table> %>%
  make_cd_equal(., code_A, code_B, direction = c("BA"))

## from code columns in data table
data %>%
  distinct(code_A, code_B) %>%
  make_cd_equal(., code_A, code_B, direction = c("AB")) %>%
  check_cd_coverage(., data, code_from, code_to) ## should definitely pass
```

## verify existing code-dict
```
code_dict %>%
  check_cd_weights(., code_from, code_to, weight_col) %>%
  check_cd_coverage(., data, code_from, code_to)
```

## apply code-dict to data
```
data_A %>%
    dplyr::group_by(case) %>%
    conformr::convert(data = .,
                    code_dict = code_dict_BA,
                    code_from = code_A,
                    code_to = code_B,
                    values_from = value_A,
                    values_to = "value_B",
                    weight_col = weight)
```

## A simple conversion
```{r simple}


```


## Complex transformations
```{r complex}
toy_AUS <- conformr:::toy_AB$data_map %>%
  dplyr::filter(country == "AUS")

## use tidy-select to convert multiple variables using the same weights
toy_AUS %>% convert(., std_A, std_B, weight, starts_with("A"))
```


## Interpreting Errors and Warnings

```{r error-warnings}
## here's a data_map I prepared earlier
toy_AB <- conformr:::toy_AB
toy_AB$data_map
```

```{r, error=TRUE}
## try convert on the data_map
toy_AB$data_map %>%
  conformr::convert(., std_A, std_B, weight, A_100)

## oops, missing groups
toy_AB$data_map %>%
  dplyr::group_by(country) %>%
  conformr::convert(., std_A, std_B, weight, A_100)

## what if weights are wrong?
toy_AB$data_map %>%
  dplyr::group_by(country) %>%
  conformr::convert(., std_A, std_B, weight_less, A_100)
```

```{r, error=TRUE}
## what if value to be split are not duplicated?
(bad_data_map <- toy_AB$data_map %>%
  dplyr::filter(country == "AUS") %>%
  dplyr::mutate(A_100 = case_when(std_B == "C3" ~ NA_real_,
                                      TRUE ~ A_100)))

bad_data_map %>%
  conformr::convert(., std_A, std_B, weight, A_100)
```
