---
title: "How to Make Panel Maps"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Making Panel Maps}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# What are Panel maps?

<!-- conceptual explainer -->

*Panel Maps* are the architectural blueprints of your combined dataset. It specifies instructions for combining and splitting values from source classification into a target classification.

# How-to make valid Panel Maps (How to Guide)

One of the key steps in the `{conformr}` workflow is the creation of *Panel Maps*. The following sections provide example code for generating *Panel Maps* from a variety of sources. Familiarity with the specific classifications, which are chosen purely for illustration purposes, is not necessary.

## Equal Weights using Correspondence Table

Say you have a correspondence table which specifies the complete connections between two code structures, how do you turn that table into a *Panel Map*?

Let's create a Panel Map to transform data from version <!-- x --> of the [Harmonised System](https://en.wikipedia.org/wiki/Harmonized_System) (`HS1_6d`) to NAICS (`code_out`) at the 6 digit code level. First, we import the relevant correspondence table, and then select the columns with the source and target codes.

```{r HS1_to_NAICS_6d_panel_map}
library(concordance)
library(magrittr)
library(dplyr)
library(conformr)

# import the correspondence table
code_dict <- concordance::hs1_naics %>%
  distinct(HS1_6d, NAICS_6d)

## "manual" way
# make a panel map, with equal weight splits
pm_df <- code_dict %>%
  group_by(HS1_6d) %>%
  mutate(n_dest = n(),
         HS1_split = 1/n_dest) %>%
  select(-n_dest)

## conformr way with de-duplication internal
# OR use the built-in to_equal_panel_map() function
panel_map <- conformr::make_panel_map_equal(code_dict, HS1_6d, NAICS_6d)

# check that it is a panel map
panel_map

# TODO: conformr::is_panel_map(HS1_to_NAICS_6d_map)
```

Notice that we kept only the unique combinations of HS1 & NAICS using `dplyr::distinct()`. We use `distinct` ensure that we have the correct denominator for splitting up values that are part of a 1-to-many split. See the [Verification of Panel Maps]() vignette for more details.

## Non-Uniform Weights

<!--# TODO: insert motivating example (e.g. BEL/LUX) -->

In simple cases: 1. Generate equal weight map 2. manually edit cell values using assignment e.g. equal[3, 7] \<- 0.3 OR 2. replace using coalesce (more tractible) uneven_weights \<- tribble(\~code_in, \~weight_new) left_join(pm, uneven_weights) %\>% coalesce(weight_new, weight)

<!--# TODO: Example of population weighted splits -->

```{r ASCO-ANZSCO}
# remotes::install_github("runapp-aus/strayr")
library(strayr)

# download correspondence table from ABS
## get(https://www.ausstats.abs.gov.au/ausstats/subscriber.nsf/0/F30E72E1516495BDCA2575DF001C7441/$File/12200%20anzsco%20first%20edition%20revision%201%20to%20asco%20second%20edition%20correspondence%20tables.xls)

## 
```

## Grouped Panel Maps

<!-- CH: Could also call this Multi-Dimensional/Hierarchical Panel Maps -->

Often when creating panel datasets, the required transformations might different between observational units. For example, <!--# TODO: add example with different map by group : e.g. isiccomb in INDSTAT4-->

## Panel Maps are Non-Invertible

Panel maps are not usually invertible. More specifically, a valid transformation map from `code_in` to `code_out` will not, in general, have the same weights as a map from `code_out` to `code_in`. One case where the panel maps are invertible are when the transformation from one classification to another involves only one-to-one mappings.

However, we can always extract the bi-directional code correspondence from an existing `panel_map` by dropping the weights column and filtering for distinct correspondences. We can then construct a new `panel_map` in the opposite direction with appropriate weights.

## TBC // From existing scripts (e.g STATA) //

In general, not possible because the panel map weights are lost at aggregation. Maybe possible from an intermediate output.

<!--# TODO: Example of extracting panel map from for loops – requires aggregation happens separately to splitting -->

``` r
# ---- make carbon paper for code/value_in ----

# scrub value_in to be 1
input_data$value_in <- 1

# ---- pass through to script ----


# ---- intermediate table **might** be panel map ----
```

## 
