---
title: "Making Valid Crossmaps"
output:
  rmarkdown::html_vignette:
    toc: yes
vignette: >
  %\VignetteIndexEntry{Making Valid Crossmaps}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Nomenclature Crossmaps

A (Nomenclature) Crossmap encodes a complete redistribution of values between a source and target classifications as a directed, bipartite, weighted graph. Numeric values transformed using a valid Crossmap will sum to the same total in both the source and target classifications. A valid Crossmap satisfies the following conditions:

1.  There is at most one link between each distinct source and target node
2.  For each source node, the sum of weights attached to outgoing links sums to one.

The outgoing and incoming properties of the graph links determine the type of data transformation encoded by a given crossmap.

1.  **Recode**: Maps with only one-to-one relations recode category labels, leaving any attached values untouched
2.  **Collapse**: Maps with at least one set of many-to-one/one-from-many links (i.e. multiple source categories collapsed into a single target category) involve some kind of aggregation of source values.
3.  **Split**: Maps with at least one set of one-to-many links (i.e. a single source category split into multiple target categories) involve redistribution of source values.

### Crossmap tables: `xmap_df`

An `xmap_df` is a data.frame representation of a Crossmap, where each row represents a weighted link between the source and target classifications. You can create an `xmap_df` from a data.frame or tibble of weighted edges using `as_xmap_df()`.

Your input table `x` needs to have at least 3 complete (i.e. no `NA`) columns:

-   `from`: source classification labels.
-   `to`: target classification labels
-   `weights`: applied to values in the source classification

Additional columns (e.g. for label descriptions etc.) can be retained using the `.keep_all` argument.

`as_xmap_df()` will validate that for `x`:

1.  There are no duplicates `from`-`to` pairs,
2.  Every group of `weights` associated with a distinct `from` value sums to 1 (subject to minor floating point errors).

## Creating Crossmaps Tables

### Row-wise creation

The following example shows how to encode one-to-one, many-to-one and one-to-many links and coerce them into a `xmap_df`. This method is most suitable for simple crossmaps.

```{r}
simple_x <- tibble::tribble(
  ~source, ~target, ~share,
  "equal", "EQUAL", 1,       # one-to-one
  "member_1", "GROUP", 1,    # many-to-one
  "member_2", "GROUP", 1,
  "whole", "PART_1", 0.3,    # one-to-many
  "whole", "PART_2", 0.6,
  "whole", "PART_3", 0.1
)

simple_xmap <- simple_x |>
  as_xmap_df(from = source, to = target, weights = share)

simple_xmap
```

### Coercion from Crosswalks

It is more common that you will want to convert an existing correspondence into a crossmap. Such conversions require attaching appropriate weights to the existing crosswalk table.

#### Recode maps

Consider the first five country codes in the ISO 3166 international standard and the one-to-one correspondence between the 2-digit, 2-digit and numeric codes.

```{r}
iso_codes <- tibble::tribble(
              ~country, ~ISO2, ~ISO3, ~ISONumeric,
         "Afghanistan",          "AF",         "AFG",    "004",
             "Albania",          "AL",         "ALB",    "008",
             "Algeria",          "DZ",         "DZA",    "012",
      "American Samoa",          "AS",         "ASM",    "016",
             "Andorra",          "AD",         "AND",    "020"
      )
```

To create a crossmap between `ISONumeric` and `ISO2`, we simply add a weights columns and coerce to `xmap_df`:

```{r}
iso_xmap <- iso_codes |>
  dplyr::mutate(link = 1) |>
  as_xmap_df(from = ISONumeric, to = ISO2, weights = link, .keep_all = TRUE)
```

Notice that `as_xmap_df()` place the `from`, `to` and `weights` columns first.

```{r}
print(iso_xmap)
```

#### Collapse maps: Aggregation

Now consider aggregating data which were collected using the ISO 3166-2 Subdivisions of [Australia](https://en.wikipedia.org/w/index.php?title=ISO_3166-2:AU&oldid=1110907059) and [Canada](https://en.wikipedia.org/w/index.php?title=ISO_3166-2:CA&oldid=1110906706):

```{r}
adm1_list <- tibble::tribble(
  ~ctr, ~adm1,
  "AU", "AU-NSW, AU-QLD, AU-SA, AU-TAS, AU-VIC, AU-WA, AU-ACT, AU-NT",
  "CA", "CA-AB, CA-BC, CA-MB, CA-NB, CA-NL, CA-NS, CA-ON, CA-PE, CA-QC, CA-SK, CA-NT, CA-NU, CA-YT"
)
```

Recall that we need one row per relation between the source (`adm1`) and target (`ctr`) nomenclature. Thus we split the string list into a vector, and then unnest the values by country.

```{r}
agg_x <- adm1_list |> 
  dplyr::mutate(adm1 = str_split(adm1, ", ")) |>
  tidyr::unnest(cols = c(adm1))

agg_x
```

Since aggregation involves the one-to-one transfer of values between `adm1` and `ctr` prior to the collapsing the `ctr` groups, we simple add weights of one to form a valid crossmap:

```{r}
agg_xmap <- agg_x |>
  dplyr::mutate(link = 1) |>
  as_xmap_df(from = adm1, to = ctr, weights = link)

#> xmap_df:
#> recode and collapse
#> (adm1 -> ctr) BY link
#> # A tibble: 21 × 3
#>    adm1   ctr    link
#>    <chr>  <chr> <dbl>
#>  1 AU-NSW AU        1
#>  2 AU-QLD AU        1
#>  3 AU-SA  AU        1
#>  4 AU-TAS AU        1
#>  5 AU-VIC AU        1
#>  6 AU-WA  AU        1
#>  7 AU-ACT AU        1
#>  8 AU-NT  AU        1
#>  9 CA-AB  CA        1
#> 10 CA-BC  CA        1
#> # … with 11 more rows
```

#### Split maps: Disaggregation

Conversely, we might have aggregate level data which want to disaggregate. Continuing the above example, this could involve incorporating country level data into analysis at the 3166-2 Subdivisions level.

For example, imagine that we have population figures for Australia at the 3166-2 level for 9 out of 10 years, but only country level figures for the missing year. In this simple example, a reasonable harmonisation design could involve splitting the country level figure by the subdivision level population proportions from the year preceding or following the missing year.

Alternatively, we might want to compare aggregate and disaggregate statistics to identify discrepancies.

```{r}
state_data <- tibble::tribble(
                              ~state,    ~adm1,    ~Pop,
                   "New South Wales", "AU-NSW", 8153600,
                          "Victoria", "AU-VIC", 6613700,
                        "Queensland", "AU-QLD", 5322100,
                   "South Australia",  "AU-SA", 1820500,
                 "Western Australia",  "AU-WA", 2785300,
                          "Tasmania", "AU-TAS",  571500,
                "Northern Territory",  "AU-NT",  250600,
      "Australian Capital Territory", "AU-ACT",  456700
)

state_xmap <- state_data |>
  dplyr::mutate(ctr = "AU",
                adm1,
                share = Pop / sum(Pop)) |>
  as_xmap_df(from = ctr, to = adm1, weights = share, .keep_all = FALSE)

#> xmap_df:
#> split
#> (ctr -> adm1) BY share
#> # A tibble: 8 × 3
#>   ctr   adm1     share
#>   <chr> <chr>    <dbl>
#> 1 AU    AU-NSW 0.314
#> 2 AU    AU-VIC 0.255
#> 3 AU    AU-QLD 0.205
#> 4 AU    AU-SA  0.0701
#> 5 AU    AU-WA  0.107
#> 6 AU    AU-TAS 0.0220
#> 7 AU    AU-NT  0.00965
#> 8 AU    AU-ACT 0.0176
```

### Splitting

Now consider the following example provided by Statistics Canada on the page [How to Read a Concordance Table](https://www.statcan.gc.ca/en/subjects/standard/concordances/concordanc_tabl3):

> In the second example, one ISIC class is equivalent to more than one NAICS Canada class. The asterisk (Partial Flag) indicates that part of ISIC D1543 is equivalent to each NAICS Canada class. The ISIC activities corresponding to each NAICS Canada class are listed in the column labelled "Link".

```{r}
canada_links <- tibble::tribble(
  ~NAICS1997, ~NAICS1997_desc, ~ISIC3, ~ISIC3_desc, ~Link,
  "311320", "Chocolate and Confectionery Manufacturing from Cacao Beans", "D1543 *", "Manufacture of cocoa, chocolate and sugar confectionery", "Chocolate and confectionery, made from cacao beans",
  "311330", "Confectionery Manufacturing from Purchased Chocolate", "D1543 *", "Manufacture of cocoa, chocolate and sugar confectionery", "Confectionery, made from purchased chocolate",
  "311340", "Non-Chocolate Confectionery Manufacturing", "D1543 *", "Manufacture of cocoa, chocolate and sugar confectionery", "Non-chocolate confectionery, manufacturing"
)

canada_links
```
