---
title: "Using Crossmaps to Transform Data"
output:
  rmarkdown::html_vignette:
    toc: yes
vignette: >
  %\VignetteIndexEntry{Using Crossmaps to Transform Data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Nomenclature Variables and Datasets

```{r echo=FALSE}
state_data <- tibble::tribble(
                              ~State,    ~`Population (est. 30 Jun 2022)`,
                   "New South Wales", 8153600,
                          "Victoria", 6613700,
                        "Queensland", 5322100,
                   "South Australia", 1820500,
                 "Western Australia", 2785300,
                          "Tasmania",  571500,
                "Northern Territory",  250600,
      "Australian Capital Territory",  456700
)

total_pop_ex <- round(sum(state_data$`Population (est. 30 Jun 2022)`) / 10^6, 2)
```

A nomenclature variable is a set of name-value pairs, where the names correspond to categories or classes in a particular nomenclature, and the values can be considered part of the same aggregate pool. For instance, consider data on population by state in Australia[^1] shown below.

[^1]: Source: [Australian Bureau of Statistics](https://www.abs.gov.au/statistics/people/population/national-state-and-territory-population/jun-2022)

```{r echo=FALSE}
knitr::kable(state_data, caption = "Australian population by State (excl. Other Territories)")
```

The set of all name-value pairs of state/territory name and population estimate for a given time period form a nomenclature variable observation for Australia. The aggregate pool for which the values for each state belong is the total population of Australia.

## Nomenclature Transformation

### Table Joins (single variable)

This vignette illustrates with simple examples how crossmaps can be used to complete recode, split and collapse transformation.

#### Aggregation

Let's consider a simple case where we have some disaggregated population figures which we want to aggregate.

First, define a crossmap for the aggregation.

```{r}
library(xmap)

agg_map <- data.frame(ctr = "AU",
                      adm1 = c("AU-NSW", "AU-QLD", "AU-SA", "AU-TAS", "AU-VIC", "AU-WA", "AU-ACT", "AU-NT"),
                      link = 1) |>
  as_xmap_df(adm1, ctr, link)
```

Second, load in our state-level figures

```{r}
state_data <- tibble::tribble(
                              ~state,    ~adm1,    ~Pop,
                   "New South Wales", "AU-NSW", 8153600,
                          "Victoria", "AU-VIC", 6613700,
                        "Queensland", "AU-QLD", 5322100,
                   "South Australia",  "AU-SA", 1820500,
                 "Western Australia",  "AU-WA", 2785300,
                          "Tasmania", "AU-TAS",  571500,
                "Northern Territory",  "AU-NT",  250600,
      "Australian Capital Territory", "AU-ACT",  456700
)
```

Now, assuming your crossmap fully covers your data the transformation involves two steps:

1.  Joining the aggregate group to each row in `state_data`
2.  Summarising values for each group

```{r}
dplyr::left_join(state_data, agg_map, by = c("adm1")) |>
  dplyr::mutate(x_pop = Pop * link) |>
  dplyr::group_by(ctr) |>
  dplyr::summarise(agg_pop = sum(x_pop))
```

Or using `apply_xmap()` from the `conformr` package:

``` r
conformr::apply_xmap(data_in = state_data, xmap = agg_map,
                     in_codes = adm1, in_values = Pop,
                     out_codes = NULL, out_values = NULL)
```

which also checks that the nomenclature in `data_in` is fully covered by your `xmap` -- i.e. you don't lose any rows during the join.

#### Disaggregation

We can reverse the aggregation (assuming we retained appropriate weights) using:

``` r
apply_xmap(data_in = ctr_data, xmap = disagg_map, in_codes = ctr, in_values = Pop, out_codes = NULL, out_values = NULL)
```

#### Complex Maps

Using crossmaps for recoding, aggregation or disaggregation can seem a bit tedious given you can achieve the same transformation with some simple join and mutate statements, coupled with some ad-hoc validation. However, crossmpas are best suited for cases where you have multiple types of transformations being applied at once.

### Matrix (multiple variables at once)
