# Core Functions

## Setup Testing Data

Define a toy example to use in development:

```{r manual-pm-equal}
## correspondence/concordance table
codes_BA <- dplyr::tribble(~ std_B, ~ std_A,
                                "A1", "x1111", # one-to-one
                                "B2", "x2222", # many-to-one
                                "B2", "x3333",
                                "C3", "x4444", # one-to-many (4)
                                "C4", "x4444",
                                "C4", "x6666", # many-to-many
                                "C5", "x4444",
                                "C6", "x4444",
                                "C7", "x5555", # one-to-many (3)
                                "C8", "x5555",
                           )

## panel_map
weights_BA <- codes_BA |>
  dplyr::distinct(std_B, std_A) |>
  dplyr::group_by(std_A) |>
  dplyr::mutate(n_dest = dplyr::n(),
                weight = 1 / n_dest) |>
  dplyr::ungroup()

pm_BA <- weights_BA |>
  dplyr::select(std_B, std_A, weight)
```

Write this data into an internal list for testing purposes.

```{r internal-toy-AB, echo=TRUE}
equal_pm <- list("codes_BA" = codes_BA,
                 "weights_BA" = weights_BA,
                 "pm_BA" = pm_BA)
```

We can visualise a panel map as the addition of weights to the concordance:

```{r viz-matrix-pm-equal}
library(ggplot2)

inc_long <- tidyr::expand(codes_BA, std_A, std_B) |>
  dplyr::left_join(pm_BA, by = c("std_A", "std_B")) |>
  dplyr::transmute(to = std_B, from = std_A, weight = weight)

gg_inc_mtx <- inc_long |> 
  plt_inc_long_mtx(to, from, weight) +
  ggtitle("Concordance as Incidence Matrix")

gg_pm_mtx <- gg_inc_mtx + 
  geom_text(data = dplyr::filter(inc_long, !is.na(weight)), aes(label=round(weight, 2))) +
  ggtitle("adding equal weights for Valid Panel Map")
```

::: {style="display: flex;"}
<div>

```{r gg-no-weight-martix}
gg_inc_mtx
```

</div>

<div>

```{r gg-equal-weight-matrix}
gg_pm_mtx
```

</div>
:::

## Valid Transformation Conditions

### Complete Mapping Weights

A valid panel map is an mapping from source to target nomenclatures which when applied to suitably dimensioned source data, transforms that data into the target nomenclature without creation or loss of value (beyond floating point rounding). This can also be thought of as a condition whereby the sum total of a variable remains the same before and after the transformation.

The following condition is necessary and sufficient for a set of Source Codes, Target Codes and Mapping Weights to be a valid panel map:

> The sum of all Mapping Weights associated with any given Source Code totals to 1

To demonstrate, let us generate some source data:

```{r}
## generate some data
set.seed(1832)
std_A_codes <- unique(codes_BA$std_A)
(data_A <- 
    dplyr::tibble(std_A = std_A_codes,
                  A_100   = 100,
                  A_prod  = round(abs(rnorm(length(std_A_codes)) * 10000),2)
                  )
  )
```

Create more testing data.

```{r internal-data-in}
equal_pm$data_A <- data_A |>
  dplyr::select(std_A, A_100)
```

Now let's switch to using the matrix representation of panel maps:

Let $\bf{C}$ be a $n \times m$ matrix showing the incidence between two disjoint sets (`inc_mtx`), and let $\bf{X}$ be the source variables (`x_mtx`) requiring transformation. Then, the transformed data is $\bf{Z} = \bf{C'X}$:

```{r pm-to-mtx}
## incidence matrix
inc_mtx <- inc_long |>
  tidyr::replace_na(list(weight=0)) |>
  inc_long_to_mtx(to, weight)

## source data matrix
x_mtx <- as.matrix(data_A[,-1])
dimnames(x_mtx)[[1]] <- std_A_codes

## transformed data
z_mtx <- t(inc_mtx) %*% x_mtx
```

::: {style="display: flex;"}
<div>

```{r print-inc-mtx}
round(t(inc_mtx), 2)
```

</div>

<div>

```{r print-data-in-mtx}
print(x_mtx)
```

</div>
:::

```{r print-data-out-mtx}
print(z_mtx)
```

Notice that the sum total of `A_100` is the same before and after the transformation.

::: {style="display: flex;"}
<div>

```{r}
colSums(x_mtx)
```

</div>

<div>

```{r}
colSums(z_mtx)
```

</div>
:::

Now, let's edit the panel map such that the weights no longer sum to one:

```{r bad-weights-pm}
## edit weights
bad_pm <- pm_BA |> 
  dplyr::mutate(weight = dplyr::case_when(
                           weight == 1 ~ weight,
                           weight < 0.5 ~ weight - 0.03,
                           weight >= 0.5 ~ weight + 0.01,
                           T ~ weight))

## incidence matrix
bad_mtx <- bad_pm |>
  inc_long_to_mtx(std_B, weight)
bad_mtx[is.na(bad_mtx)] <- 0

## transform data badly
bad_z <- t(bad_mtx) %*% x_mtx
```

Notice what happens when we apply the transformation:

::: {style="display: flex;"}
<div>

```{r print-bad-mtx}
round(t(bad_mtx), 2)
```

</div>

<div>

```{r print-bad-data-in}
print(x_mtx)
```

</div>
:::

```{r print-bad-data-out}
print(bad_z)
```

Notice that the sum totals are no longer the same before and after the transformation:

::: {style="display: flex;"}
<div>

```{r}
colSums(x_mtx)
```

</div>

<div>

```{r}
colSums(bad_z)
```

</div>
:::

Hence, the validity condition can also be expressed as follows: \> A given incidence matrix $\bf{K}$ with dimensions $n \times m$ is a valid panel map if and only if $\bf{K}\boldsymbol{1} = \boldsymbol{1}$ where $\boldsymbol{1}$ is a unit vector of length $m$:

```{r valid-weights-mtx}
ones <- rep_len(1, ncol(inc_mtx))
```

::: {style="display: flex;"}
<div>

```{r}
round(inc_mtx, 2)
```

</div>

<div>

```{r}
inc_mtx %*% ones
```

</div>
:::

#### Functions

Internal switching function for flow control and error messages

```{r fnc-flag-weights}
#' Flag Bad Mapping Weights
#' 
has_bad_weights <- function(df, code_in, code_out, weights){
  bad_rows <- df |>
    dplyr::group_by({{code_in}}) |>
    dplyr::summarise(total = sum({{weights}}),
                     weights = paste({{weights}}, collapse=",")) |>
    dplyr::filter(total != 1)
  
  is_bad <- !(nrow(bad_rows) == 0)

  result <- list(fail = is_bad,
                 table = bad_rows)

  return(result)
}
```

This function checks if the panel map has valid weights and returns the panel map if it does. It can be used to validate a panel map after editing or modifications. For example:

``` r
## prepare panel map
new_pm <- old_pm |>
  mutate() |>
  filter() |>
  check_pm_weights(code_in, code_out, weights)
```

```{r fnc-check-weights}
#' Check panel map weights are valid
#' 
#' Checks if `code_in`, `code_out` and `weights` columns of data frame forms a valid panel map.
#' 
#' @param df Data Frame containing weighted links `weights` between `code_in` and `code_out`.
#' @param code_in Variable in `code_dict` containing source codes to convert from.
#' @param code_out Variable in `code_dict` containing destination codes to convert to.
#' @param weights Column containing weights for transforming values from `code_in` to `code_out`
#'
#' @exports
#' 
#' @returns The original data frame if the check is passed and an error if not.
check_weights <- function(df, code_in, code_out, weights){
  
  has_result <- has_bad_weights(df, {{code_in}}, {{code_out}}, {{weights}})
  
  if (has_result$fail){
    
    cli::cli_abort(c(
        "{.var weights} for each {.var code_in} must sum to 1",
        ""
        ),
        class="invalid_weights"
        )
  } else {
    return(df)
  }
}
```

#### Tests

-   flag function returns expected output
-   check function works as expected:
    -   returns informative error message
    -   returns unchanged panel map

Add testing data

```{r internal-test-weights}
equal_pm$bad_weights <- equal_pm$pm_BA |>
  dplyr::mutate(weight = dplyr::case_when(
                           weight == 1 ~ weight,
                           weight < 0.5 ~ weight - 0.03,
                           weight >= 0.5 ~ weight + 0.01,
                           T ~ weight))
```

Write tests:

```{r testthat-weight-fncs}
testthat::test_that(
  "has_bad_weights() returns correct flags",
  {
    # good weights
    testthat::expect_false(
      has_bad_weights(equal_pm$pm_BA, std_A, std_B, weight)$fail
      )
    # bad weights
    testthat::expect_true(
      has_bad_weights(equal_pm$bad_weights, std_A, std_B, weight)$fail
      )
  }
)
testthat::test_that(
  "check_weights() works as expected",
  {
    # good weights
    testthat::expect_identical(
      check_weights(equal_pm$pm_BA, std_A, std_B, weight), equal_pm$pm_BA)

    # bad weights
    testthat::expect_error(
      check_weights(equal_pm$bad_weights, std_A, std_B, weight),
      class="invalid_weights"
      )
  }
)
```

### No Missing Data Values

Except for a one-to-one transfer between classifications, there is no way for NA values in the Source Data to be preserved when transformed into the Target Classification. It doesn't make sense to split NA into smaller parts, or to aggregate NA into a sum.

Hence, any missing values need to be explicitly dealt with before applying a Panel Map transformation. Exactly how missing values should be treated will vary from dataset to dataset. This could involve replace the missing values with zeroes or some imputed values, or to remove them completely.

```{r gg-na-data-in}
pm_BA |>
  plt_pm_sigmoid(from=std_A, to=std_B, weights = weight) +
  scale_fill_brewer(palette="RdPu", direction=-1)
```

#### Functions

This function flags if the variables you want to transform have any missing values.

```{r fnc-flag-na}
#' Flags NA in Source Data
#'
has_missing <- function(.data){
  is_miss <- .data |>
    anyNA()

  result <- list(fail=is_miss)

  return(result)
}
```

This function checks the dataframe for missing values, and returns the original dataframe or tells the user to fix the NAs in their data. The dataframe should already be subsetted to contain only the Source Code and Source Value columns:

``` r
## prepare data for transformation
data_in <- all_df |>
  select(code_in, x1, x2) |>
  check_missing()
```

```{r fnc-check-na}
#' Checks Source Data for Missing Values
#'
#' @inheritParams concord
#'
#' @export
check_missing <- function(data_in){
  has_result <- has_missing(data_in)
  
  if(has_result$fail){
    cli::cli_abort(
      "{.var data_in} should not have any NA",
      class="vals_na"
    )
  } else {
    return(data_in)
  }
}
```

#### Tests

Feed in data with missing values and expect: - TRUE flag - Error message

Add testing data

```{r internal-data-bad}
equal_pm$bad_data <- equal_pm$data_A
equal_pm$bad_data[1, 2] <- NA
```

```{r testthat-na}
testthat::test_that(
  "has_missing() returns expected flags",
  {
    # good weights
    testthat::expect_false(
      has_missing(equal_pm$data_A)$fail
      )
    
    # bad weights
    testthat::expect_true(
      has_missing(equal_pm$bad_data)$fail
      )
  }
)
testthat::test_that(
  "check_missing() works as expected",
  {
    ## good data
    testthat::expect_identical(check_missing(equal_pm$data_A), equal_pm$data_A)
    
    ## bad data
    testthat::expect_error(check_missing(equal_pm$bad_data),
                           class = "vals_na")
  }
)
```

### Source Code Coverage

A Panel Map must cover all Source Codes present in the Source Data. In other words, for a transformation to be valid, no Source Data should be left behind.

```{r gg-data-in-mtx}
gg_x_mtx <- plt_df_mtx(data_A, A_100:A_prod, std_A)

library(patchwork)
gg_pm_mtx + 
  guides(fill="none") +
  ggtitle("") + gg_x_mtx +
  scale_y_discrete(position="right", limits=rev) +
patchwork::plot_annotation(title="Panel Map covers Source Data")
```

```{r gg-no-coverage}
gg_x_bad <- data_A |>
  dplyr::add_row(std_A = "x7285!", 
          A_100 = 100, 
          A_prod = 3895.3) |>
  plt_df_mtx(A_100:A_prod, std_A)

gg_pm_bad <- tidyr::expand_grid(from=c(NA), to=unique(codes_BA$std_B)) |>
  dplyr::mutate(weight=NA) |>
  bind_rows(inc_long) |>
  plt_inc_long_mtx(to, from, weight) +
  geom_text(data = dplyr::filter(inc_long, !is.na(weight)), aes(label=round(weight, 2)))


library(patchwork)
gg_pm_bad + 
  guides(fill="none") +
  ggtitle("") + 
gg_x_bad +
  scale_y_discrete(position="right", limits=rev) +
  scale_fill_brewer(palette="Purples") +
patchwork::plot_annotation(title="Panel Map does not cover fully Source Data")
```

Depending on how the transformation is implemented, coverage mismatches can result in both explicit and implicit/hidden errors. In particular, having conformable matrix dimensions is not sufficient to avoid corrupting data unless you check that the indices match. This is a common issue with using matrices for data wrangling, so this package implements transformations using database operations.

#### Functions

Internal checking function -- assumes `.map` is a valid map. Note this could be (quickly) checked for using a class condition discussed in [GitHub issue #43](https://github.com/cynthiahqy/conformr-project/issues/43)

```{r fnc-coverage-flag}
#' Flag if data set is not completely cover by panel map
#' 
#' @inheritParams use_panel_map
#'  
has_coverage <- function(.data, .map, .from){
  
  missing_links <- .data |>
    dplyr::select(tidyselect::all_of(.from)) |>
    dplyr::distinct() |>
    dplyr::anti_join(.map, by = .from)

  is_covered <- (nrow(missing_links) == 0)

  results <- list(fail=!is_covered,
                  table=missing_links)

  return(results)
}
```

Error constructing function, also used in `concord()`
```{r fnc-coverage-check}
#' Check coverage of panel map over source data
#' 
#' @inheritParams concord
#' @inheritParams use_panel_map
#' 
#' @returns `data_in` if check is successful, throws error otherwise.
#' @examples
#' 
#' /notrun{
#' check_coverage(df, pm, "std_A")
#' }
#' 
#' 
check_coverage <- function(data_in, pm, .from){
  # call flag function
  has_result <- has_coverage(data_in, pm, .from)
  
  # conditionals
  if(has_result$fail){
    cli::cli_abort(
      "{.var data_in$from_code} has values not covered by {.var pm$from_code}",
      class="not_covered"
    )
  } else {
    return(data_in)
  }
  
}
```

#### Tests

Add some more testing data

```{r internal-data-not-covered}
equal_pm$data_extra <- equal_pm$data_A |>
      dplyr::add_row(std_A = "x7777", A_100 = 100)
```

```{r testthat-coverage-flag}
testthat::test_that(
  "has_coverage() returns expected flags",
  {
    ## complete coverage
    testthat::expect_false(has_coverage(equal_pm$data_A, equal_pm$pm_BA, "std_A")$fail)
    
    ## incomplete coverage
    testthat::expect_true(has_coverage(equal_pm$data_extra, equal_pm$pm_BA, "std_A")$fail)
  }
)
```


```{r testthat-coverage-check}
testthat::test_that(
  "check_coverage() works as expected",
  {
    ## complete coverage
    testthat::expect_identical(check_coverage(equal_pm$data_A, equal_pm$pm_BA, "std_A"), equal_pm$data_A)
    ## incomplete coverage
    testthat::expect_error(check_coverage(equal_pm$data_extra, equal_pm$pm_BA, "std_A"),
                           class = "not_covered")
  }
)
```

## Use Panel Map on Data

### Single Step Concordance

#### Stylized Code

Assuming all the validity conditions are met, we want a simple and concise way to apply a panel map to data which looks something like:

``` r
# --- prepare panel map --------------------
df_pm <- read_csv("concordance-table.csv") |>
  conformr::make_panel_map_equal(...) |>
  conformr::validate_panel_map(...)
  
# --- prepare data ---------------------------
df_data_in <-
  read_csv("your-source-data.csv") |>
  conformr::validate_data_in(...)
  
## --- apply (valid) transformation -----------
conformr::concord(
  data_in = df_data_in, pm = df_pm,
  from_code = source, to_code = target,
  m_weights = weight, values_from = value_in,
  .suffix = "_out"
)
```

Preparing a panel map and data for valid transformation could look like:

``` r
## --- prepare panel map -------------------- ##
# by importing a manually encoded map
df_pm <- read_csv("your-panel-map.csv")
# or creating one from a concordance table
df_pm <- read_csv("concordance-table.csv") |>
  conformr::make_panel_map_equal(
    code_in = source, code_out = target,
    .weights_to = "weight")

## --- prepare source data ------------------ ##
# example using {dplyr}:
df_data_in <-
  read_csv("your-source-data.csv") |>
  drop_na() |>
  group_by(source) |>
  summarise(value_in = sum(gdp))

## --- apply (valid) transformation --------- ##
conformr::concord(
  data_in = df_data_in, pm = df_pm,
  from_code = source, to_code = target,
  m_weights = weight, values_from = value_in,
  .suffix = "_out"
)
```

#### Warnings and Errors

The concordance function should throw error when:

-   panel map (`pm`) has invalid weights
-   source data (`data_in`) column has missing values

The concordance function should warn users about data prep?:

-   multiple rows for a given `code_in` in `data_in`; should only have one set of `value_in` for each `code_in`

#### Functions

This function takes a valid panel map and data with matching names for the Source Code columns and transforms the data to the Target Classification.

Add informative error messages later:

``` r
  in_data_in <- (str.vals %in% colnames(data_in))
  if (!all(in_data_in)){
    cli::cli_abort(
      "{.code {names(dots)[!in_data_in]}} cannot be found in {.var data_in}",
      class = "cols_not_found")
  }
```

```{r fnc-concord}
#' Transform data from Source to Target classification using Panel Map
#' 
#' Currently checks for valid Mapping weights, missing values, and coverage.
#'
#' @param data_in A Data Frame containing the values you want to transform
#' @param pm A Data Frame containing valid Mapping Weights between `from_code` and `to_code`.
#' @param from_code Variable containing Source Codes. Must be present in both `data_in` and `pm`
#' @param to_code Variable in `pm` containing Target Codes.
#' @param m_weights Variable in `pm` containing Mapping Weights.
#' @param values_from A vector of variables in `data_in` to be transformed. E.g. `c(var1, var2)`
#' @param .suffix An (optional) string appended to each `values_from` name to create column names for transformed values.
#' Defaults to `"_out"`
#' 
#' @return
#' @export
#'
#' @examples
#' \dontrun{
#' concord(data_in = equal_pm$data_A,
#'        pm = equal_pm$pm_BA,
#'        from_code = std_A,
#'        to_code = std_B,
#'        m_weights = weight,
#'        values_from = c(A_100),
#'        .suffix = "_out")
#' }
#' 
concord <- function(data_in, pm, from_code, to_code, m_weights, values_from, .suffix=NULL){
  
  ## defuse arugments
  str.to <- rlang::as_string(rlang::enexpr(to_code))
  str.from <- rlang::as_string(rlang::enexpr(from_code))
  
  ## check conditions
  pm |> 
    check_weights(code_in = {{from_code}},
                  code_out = {{to_code}},
                  weights = {{m_weights}})
  
  subset_in <- tryCatch(
    data_in |>
    dplyr::select({{from_code}}, {{values_from}}), 
    error = function(cnd) {
      cli::cli_abort(
      "{.var from_code} or {.var values_from} could not be found in {.var data_in}",
      class = "vals_not_found")
    }
  )
  
  subset_in |>
    check_missing()
  
  check_coverage(subset_in, pm, str.from)
  
  ## apply transformation
  # -- create suffix --
  out_suffix <- .suffix %||% paste0("_", str.to)
  join_by <- str.from
  
  data_out <- use_panel_map(.data = subset_in, .map = pm, 
                .from = {{from_code}}, .to = {{to_code}}, .weights = {{m_weights}},
                .vals = {{values_from}}, .suffix = out_suffix,
                .by = join_by)
  
  return(data_out)
}
```

Internal function without checks

```{r fnc-use-panel-map}
#' Apply panel_map to data without checks
#'
#' A wrapper around a `{dplyr}` pipeline that takes a panel_map,
#' joins it with data, and transforms selected variables in that data according to
#' instructions in the panel map. Any groups in `data_in` are preserved.
#'
#' @param .data a Data Frame assumed to meet Source Data conditions
#' @param .map a Data Frame assumed to meet Panel Map conditions
#'
#' @return The output has the following properties:
#' * Groups are taken from `data_in`
#'
use_panel_map <- function(.data, .map, .from, .to, .weights, .vals,
                          .suffix, .by){
  
  # subset data for transformation
  data_in <- .data %>%
    dplyr::select({{.from}}, {{.vals}})

  # merge map and data // use default by= argument
  map_join_data <- dplyr::right_join(x = data_in,
                                     y = .map,
                                     by = .by)

  # apply transformation
  data_out <- map_join_data %>%
    dplyr::mutate(dplyr::across({{ .vals }}, ~ .x * {{ .weights }})) %>%
    dplyr::group_by({{ .to }}, .add = TRUE) %>%
    dplyr::summarise(dplyr::across({{ .vals }}, ~ sum(.x)), .groups = "drop_last")

  # rename
  data_out <- data_out %>%
    dplyr::rename_with(., ~ paste0(.x, .suffix), .cols = {{.vals}})

  return(data_out)
}
```

#### Tests

Define some test data:

```{r internal-data-out}
equal_pm$data_B <- 
  dplyr::right_join(x = equal_pm$data_A,
                    y = equal_pm$pm_BA,
                    by = "std_A") |>
  dplyr::mutate(A_100 = A_100 * weight) |>
  dplyr::group_by(std_B, .add = TRUE) |>
  dplyr::summarise(dplyr::across(c(A_100), ~ sum(.x), .names = "{.col}_out"),
                   .groups = "drop_last")
```

Do the tests:

```{r testthat-use-pm}
testthat::test_that(
  "use_panel_map() works as expected", {
    testthat::expect_identical(
      use_panel_map(.data = equal_pm$data_A,
              .map = equal_pm$pm_BA,
              .from = std_A,
              .to = std_B,
              .weights = weight,
              .vals = c(A_100),
              .suffix = "_out",
              .by = "std_A"),
      equal_pm$data_B
    )
  }
)
```

```{r testthat-concord-error}
testthat::test_that(
  "concord() raises expected errors",
  {
    ## columns not in data_in
    testthat::expect_error(concord(data_in = equal_pm$data_A,
                                   pm = equal_pm$pm_BA,
                                   from_code = std_A,
                                   to_code = std_B,
                                   m_weights = weight,
                                   values_from = c(missing_col1, missing_col2)
                                   ),
                           class="vals_not_found")
    ## missing values in data_in
    testthat::expect_error(concord(equal_pm$bad_data, equal_pm$pm_BA, std_A, std_B, weight,
                                   values_from = c(A_100)
                                   ),
                           class="vals_na"
                           )
    ## invalid weights are flagged
    testthat::expect_error(concord(equal_pm$data_A, equal_pm$bad_weights, std_A, std_B, weight,
                                   values_from = c(A_100)
                                   ),
                           class="invalid_weights"
                           )
  }
)
```

```{r testthat-concord-pass}
testthat::test_that(
  "concord() works as expected",
  {
    testthat::expect_identical(concord(data_in = equal_pm$data_A,
                             pm = equal_pm$pm_BA,
                             from_code = std_A,
                             to_code = std_B,
                             m_weights = weight,
                             values_from = c(A_100),
                             .suffix = "_out"),
                     equal_pm$data_B
                     )
  }
)
```
